Deciding Bitvector Termination with Program Synthesis

### 2014-06-25 papers/termination/paper.tex
Heap is a limitation
added non-termination to related work

\todo{add non-termination}

@@

%\subsection{Synthesis}

@@

%In the previous sections, we have discussed some of the assumptions made by existent termination analyses (see Figure~\ref{fig:handletable}). 
Next, we show how these assumptions are broken by even simple programs that can't therefore be handled accurately by state of the art termination tools.

@@

Figure~\ref{fig:motivation.g} is a linear program taken from \cite{DBLP:conf/tacas/CookSZ13}, 
where it is shown to not admit (without prior manipulation) a lexicographic linear function, but only a Ramsey-based termination argument,
which requires an expensive binary reachability analysis.
With our technique we can find a non-linear lexicographic ranking function $R(x) = ...$.
A lexicographic non-linear ranking functions consists of lexicographically ordered components
of non-linear functions. 
%As with linear lexicographic ranking functions, a state is mapped to a tuple of values such that the
%loop transition leads to a decrease with respect to the lexicographic
%ordering for this tuple. Therefore no function may increase unless a function of
%a lower index decreases. Additionally, at every step, there must be at least one
%function that decreases.

@@

The loop in Figure~\ref{fig:motivation.h} illustrates conditional termination.
Basically, when proving program termination we are simultaneously solving two problems:
the search for a termination argument, and the search for a supporting invariant.
We iteratively construct a termination argument as a pair of a ranking function and supporting invariant through a single refinement loop.
For this loop, we find the ranking function $R(x) = x$ together with the supporting invariant $y=1$.

>>>>

@@

While program termination has been extensively studied, much less research has been conducted in the area of proving nontermination.
% A counterexample to termination is an infinite program execution. 
Gupta et al. dynamically enumerate lasso-shaped candidate
paths for counterexamples, and then statically prove their feasibility \cite{DBLP:conf/popl/GuptaHMRX08}. 
In \cite{DBLP:conf/tacas/ChenCFNO14}, Chen et al. prove nontermination via
reduction to safety proving. Their iterative algorithm uses counterexamples to a fixed safety property
to refine an underapproximation of a program.

@@

%In the previous sections, we have discussed some of the assumptions made by existent termination analyses (see Figure~\ref{fig:handletable}). 
Next, we show how these assumptions are broken by even simple programs that cannot therefore be handled accurately by state of the art termination tools.

@@

Figure~\ref{fig:motivation.g} is a linear program taken from \cite{DBLP:conf/tacas/CookSZ13}, 
where it is shown to not admit (without prior manipulation) a lexicographic linear function, but only a Ramsey-based termination argument,
which requires an expensive binary reachability analysis.
With our technique we can find a non-linear lexicographic ranking function $R(x) = |x|$ 
(a lexicographic non-linear ranking functions consists of lexicographically ordered components
of non-linear functions). 
%As with linear lexicographic ranking functions, a state is mapped to a tuple of values such that the
%loop transition leads to a decrease with respect to the lexicographic
%ordering for this tuple. Therefore no function may increase unless a function of
%a lower index decreases. Additionally, at every step, there must be at least one
%function that decreases.

@@

The loop in Figure~\ref{fig:motivation.h} illustrates conditional termination.
When proving program termination we are simultaneously solving two problems:
the search for a termination argument, and the search for a supporting invariant.
We iteratively construct a termination argument as a pair of a ranking function and supporting invariant through a single refinement loop.
For this loop, we find the ranking function $R(x) = x$ together with the supporting invariant $y=1$.
